---
title: "Practical Machine Learning Assignment"
output: html_document
---

# Summary

A random tree model was trained using the provided data. The resultant model had an estimated out of sample error of <0.2%. When applied to the testset provided, the model was able to accurately identify all 20 activities listed.

##Setting up environment and loading relevant data

```{r message=FALSE}
set.seed(11)
library(caret)
library(doParallel)
registerDoParallel(cores=3)

testset<-read.csv("pml-testing.csv",header=TRUE)
trainset<-read.csv("pml-training.csv",header=TRUE)

```

##Pre-processing the data and segregation into training and test sets

```{r}
#Remove predictors that are not present in the final test set
allna <- sapply(testset, function(x) mean(is.na(x))) ==1
trainset <- trainset[, allna==F]
testset <- testset[, allna==F]

#Removing predictors with near zero variation
nzvar <- nearZeroVar(trainset)
trainset <- trainset[, -nzvar]
testset <- testset[, -nzvar]

#Removing predictors that have mostly NA values
almost_NA <- sapply(trainset, function(x) mean(is.na(x))) > 0.95
trainset <- trainset[, almost_NA==F]
testset <- testset[, almost_NA==F]

#Removing irrelevant predictors 
trainset <- trainset[, -c(1,3,4,5)]
testset <- testset[, -c(1,3,4,5)]

#Partitioning into test and training sets
inTrain<-createDataPartition(y=trainset$classe, p=0.7, list=FALSE)
inTraining<-trainset[inTrain,]
inTesting<-trainset[-inTrain,]


```


##Training and validating the model

```{r cache=TRUE, message=FALSE}
tcterm<-trainControl(method = "boot", number=5,allowParallel = TRUE)
modFit<-train(classe~.,data=inTraining, method="rf",prox=TRUE, trControl = tcterm)
modFit$finalModel

predictval<-predict(modFit,newdata=inTesting)
confusionMatrix(predictval,inTesting$classe)
```

Performance on the training set was satisfactory, with 99.77% of entries correctly classified.
Cross validation results in an accuracy of 99.86%, suggesting that the out of sample error can be estimated less than 0.2%.

##Re-training the model
Now that the model has been validated, we can re-train it using the full set of data.

```{r message=FALSE, cache.lazy=FALSE}
finalFit<-train(classe~.,data=trainset, method="rf",prox=TRUE, trControl = tcterm)
finalFit$finalModel
```



##Applying the model to the test set
Applying the model to the test set gives the results as below:
```{r cache=TRUE}
testval<-predict(finalFit,newdata=testset)
testval
```

